Methods for Analyzing Models of\\ Rod-Shaped Bacteria

### 2025-09-20 paper/main.tex
rewrite individual treatment section

% --------------------------------------------------------------------------------------------------
\subsection{Individual Treatment}
We have already suspected (see table~\ref{table:simulation-aspects}) that in order to properly
estimate parameters of our system, we need to treat every cell-agent individually, thus assigning
parameters individually to them.
This means that we increase the total number of parameters that needs to be sampled within an
optimization routine.
Given $N_c$ cells and $N_p$ parameters per cell and $N_u$ unique parameters (where $N_u\leq N_p$),
the total amount of parameters is

@@

It is clear, that we should avoid to needlessly assign every parameter individually since the space
which is needed to be sampled scales exponentially with the dimension ie. the number of parameters.
However, we will now present data that indicate that an individual treatment is necessary.
Figure~\ref{fig:estimated-growth-rates} shows the rod-lengths of individual bacteria which have been
calculated from the extracted positions.
We fit a delayed exponential function to them which is defined by

@@

and where $\lambda$ is the growth rate $t_0$ the delay.
The fits to invidiual growth-curves of figure~\ref{fig:estimated-growth-rates} (A) result in good
matches and so does the fit of subfigure (B).
However, the latter exhibits large amounts of uncertainty within the data which is also illustrated
in subfigure (C).
There, we can see much more clearly the difference between considering invidiually assigned
parameters and global ones.
It is clear from this observation that in order for us to accurately estimate the parameters of this
system, we need to treat the bacteria individually.

The individual treatment of the bacteria allows us in principle to estimate the distribution of the
parameters directly without having to resort to any assumptions on its form.
This is true for any parameter, not just the growth rates, since we aim to estimate all mechanical
parameters of our cellular model.

\textbf{TODO is this discussion?}\\
The total parameter space $P$ is a product of all single spaces $P_i$ in which the parameters
$p_i$ live $P=P_0\times P_1\times\dots$ but some of these parameters share the same space (let's
assume $p_0,p_1\in \tilde{P}$).
This leaves the open question as to weather the optimization algorithm could be modified to leverage
this knowledge and obtain increased performance.
For instance, it might be valuable to initially consider individual parameters as globally defined
and start treating the individually after a certain amount of optimization steps.

>>>>

% --------------------------------------------------------------------------------------------------
\subsection{Individual Treatment}
We have already suspected (see table~\ref{table:simulation-aspects}) that in order to properly
estimate parameters of our system, we need to treat every cell-agent individually, thus assigning
parameters on a single-cell level.
This would mean to increase the total number of parameters that need to be sampled within an
optimization routine.
Given $N_c$ cells and $N_p$ parameters per cell and $N_u$ individual parameters (where
$N_u\leq N_p$), the total amount of parameters is

@@

It is clear, that we should avoid to needlessly assign every parameter individually since the volume
of the space that needs to be sampled, scales exponentially with the dimension ie. the number of
parameters.
However, we will now present data that indicate that an individual treatment is in fact necessary.
Figure~\ref{fig:estimated-growth-rates} shows the rod-lengths of individual bacteria which have been
calculated from the extracted positions.
We fit a delayed exponential function to them which is defined by

@@

where $\lambda$ is the growth rate and $t_0$ the delay.
Fitting this function to the individual growth curves of the bacteria and their mean growth curve,
both results in good agreements (Figure~\ref{fig:estimated-growth-rates} A,B).
However, the latter exhibits large amounts of uncertainty within the data which is also illustrated
in subfigure (C).
There, we can see much more clearly the difference between considering invidiually assigned
parameters and global ones.
It is clear from this observation that in order for us to accurately estimate the parameters of this
system, we need to treat the bacteria individually.
